Sediment transport capacity response to variations in water discharge in pressurized subglacial channels

### 2025-02-03 revised_v1/response_R1.tex

\documentclass[11pt]{article}
\usepackage{lscape}
\usepackage[table]{xcolor}% http://ctan.org/pkg/xcolor
\setlength{\parindent}{0pt}
\usepackage [colon]{natbib}
\usepackage {comment}
\renewcommand{\familydefault}{\sfdefault}
\usepackage[a4paper, margin=2.40cm]{geometry}
\usepackage{doi}
\usepackage{multirow}
\usepackage[T1, OT1]{fontenc}
\DeclareTextSymbolDefault{\dh}{T1}
\renewcommand{\rmdefault}{phv} % Arial
\renewcommand{\sfdefault}{phv} % Arial
\usepackage[modulo]{lineno}
\usepackage{wrapfig}
% \linenumbers
\usepackage{hyperref}
\usepackage{array}
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\usepackage{amsmath}
\usepackage{float}
% \errorcontextlines=3
\usepackage[explicit]{titlesec}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{authblk}
\usepackage{setspace}
\usepackage[export]{adjustbox}
\renewcommand{\bibfont}{\footnotesize}
\usepackage{bibentry}
\usepackage{tikz}
\usepackage[inline,shortlabels]{enumitem}
\usepackage[all]{nowidow}

\newcommand{\TODO}[1]{\textbf{TODO #1}}
\newcommand{\ian}[1]{{\textbf{\color{blue}Ian says:} \color{blue} #1} }
\newcommand{\leif}[1]{{\textbf{\color{red}Leif says:} \color{red} #1} }
\newcommand{\frederic}[1]{{\textbf{\color{red}Frederic says:} \color{red} #1} }
\newcommand{\aref}[1]{\textbf{Reference #1}}
\newcommand{\alpine}{\textit{ALPINE}\,}
\newcommand{\icesheet}{\textit{ICESHEET}\,}
\title{}
\author{}

\begin{document}
Dear Editor,

\vspace{.3cm}

We thank Reviewer 1 for their detailed and constructive comments. There work, we believe, has provided feed back that has greatly improved the manuscript.

The reviewer's comments are in bold, our response is in italics and quotations from the new text are in normal font.

\vspace{.3cm}

@@

\item \textbf{171-} I get a little muddled right at the start here – not helped by the sub-heading,
  which I would recast as “Influence of subglacial channel size on the timing
  and variability of…”. The section (and others following) also combines
  results with explanation (i.e., interpretation) which I feel also adds
  unnecessarily to the difficulty in following the progression. I would separate
  results from interpretation.

  \textit{This section will be retitled:} ``Timing of subaerial and subglacial sediment transport capacity variations.''
  \textit{This discusses the role water discharge on hysteresis in water velocity or sediment transport capacity. Thus dealing with objective 1 of the manuscript. We have also noted the comment about results and interpretation. These have been removed from the manuscript}

\item \textbf{172} This narrative jumps straight in with the ambition to ‘quantify the sources
  of increased variability…’ but the nature of that increased variability has not
  yet be established or illustrated. This really needs building up more
  progressively – bringing the reader through with the first-order model
  results. Then go on to explore more and more detailed influences and
  relationships. Most of this is, in fact, in Figure 2 – but the reader is not
  guided through the fundamental relationships here.

 \textit{Excellent comment. The first sentence of section will read:} ``The first numerical experiment aims to quantify the timing and covariance of the subglacial model outputs with respect to water discharge in response to different seasonal evolutions and peaks''

@@

  \textit{We believe that description of the insets' axis is needed as these lines have the same axis in the main plots. However, their very different values mean that in makes sense to adjust values to examine their changes during the extreme events. The text has been modified slightly to read: ``different y-axis ranges for subaerial and subglacial values.''}

\item \textbf{179-183} Is this not already established? If so, it should be presented clearly in the
  Introduction and developed as necessary here. It is also interpretation.

  \textit{ To the best of our knowledge this is the first time that this finding is being presented.
    We would welcome any citations or references that suggest otherwise. We believe that the content in this paragraph describes the response to water discharge in Figure 2. We agree that the reference to the Methods section, gave the impression of the interpretation. The paragraph will be adjusted slightly to read:}

    Peaks in subaerial model outputs occur coincident with peaks in water discharge (Figure~2).
    In the subglacial channel, peaks in model outputs generally occur when water discharge increases, but before the maximum water discharge.
    As the water discharge stabilizes at its peak, channel growth continues (Figure~2 a, e), causing water velocity and other model outputs to decrease from their peak values (Figure~2 b-d, f-h).
    Subglacial sediment transport capacity is greatest on the hydrograph's rising limb, relative to the falling limb, creating a hysteresis effect.''

### 2025-02-03 revised_v1/response_R2.tex

\documentclass[11pt]{article}
\usepackage{lscape}
\usepackage[table]{xcolor}% http://ctan.org/pkg/xcolor
\setlength{\parindent}{0pt}
\usepackage [colon]{natbib}
\usepackage {comment}
\renewcommand{\familydefault}{\sfdefault}
\usepackage[a4paper, margin=2.40cm]{geometry}
\usepackage{doi}
\usepackage{multirow}
\usepackage[T1, OT1]{fontenc}
\DeclareTextSymbolDefault{\dh}{T1}
\renewcommand{\rmdefault}{phv} % Arial
\renewcommand{\sfdefault}{phv} % Arial
\usepackage[modulo]{lineno}
\usepackage{wrapfig}
% \linenumbers
\usepackage{hyperref}
\usepackage{array}
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\usepackage{amsmath}
\usepackage{float}
% \errorcontextlines=3
\usepackage[explicit]{titlesec}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{authblk}
\usepackage{setspace}
\usepackage[export]{adjustbox}
\renewcommand{\bibfont}{\footnotesize}
\usepackage{bibentry}
\usepackage{tikz}
\usepackage[inline,shortlabels]{enumitem}
\usepackage[all]{nowidow}

\newcommand{\alpine}{\textit{ALPINE}\,}
\newcommand{\icesheet}{\textit{ICESHEET}\,}
\newcommand{\TODO}[1]{\textbf{TODO #1}}
\newcommand{\ian}[1]{{\textbf{\color{blue}Ian says:} \color{blue} #1} }
\newcommand{\leif}[1]{{\textbf{\color{red}Leif says:} \color{red} #1} }
\newcommand{\frederic}[1]{{\textbf{\color{red}Frederic says:} \color{red} #1} }
\newcommand{\aref}[1]{\textbf{Reference #1}}
\title{}
\author{}

\begin{document}
Dear Editor,

\vspace{.3cm}

We thank Reviewer 2 for their comments. We believe that in addressing them, we have clarified the model presentation. We note, however, that the presentations of criticisms by the reviewer was at times difficult to respond to, given the format of the review. Please let us know if additionally points should be addressed.

The reviewer's comments are in bold, our response is in italics and quotations from the new text are in normal font.

\vspace{.3cm}

Best regards,

\vspace{.75cm}

Ian Delaney on behalf of all authors

\vspace{2cm}

@@

\textbf{General Comments}

\begin{itemize}

\item \textbf{
    The trouble is, that while the first term on the right hand side of (3) makes sense to me, the second does not, and the reason for this is that the closure term involves the effective pressure and thus the hydraulic head, but not its gradient, so I am a bit sceptical about this, because the structure of the Nye model seems to have been altered.
    You might say that the distinction disappears in the lumping, but it seems to me that if you have an upstream head h- and a downstream head h+, the first term on the right of (3) will involve the difference, but the second will involve the average, and there are two independent quantities involved.
    I suppose you can get around this by saying, oh, the outlet head or more accurately the effective pressure is zero, and perhaps that is what is done.
    But that is a bit disingenuous, because the outflow becomes open channel flow somewhere upstream of the actual mouth of the stream.
    Now I went and looked up this Werder 2010 paper, and you can see in its equation (4) the same distinction I am making here.
    I suppose my overall view is that (I think) the Clarke paper intended to use the flow elements of resistors, etc., as ingredients of a larger scale flow path, and this lumping of a whole channel is an extremely coarse thing to do, and looks a bit like avoiding confronting the spatially-dependent physics simply because it’s too hard.
    So I think you can rescue this aspect of the model presentation, but a bit more exposition would help, in particular the figure 2or equivalent of the 2010 paper would help.}

  \textit{We thank the reviewer for this comment, as it demands clarification in our model presentation. In order to reduce the number of variables in the original text, $h_p$ was omitted as its value is zero, as is hydraulic head at the glacier terminus. The text has been updated to clarify this point.
    Additionally, we confirm the reviewers comment, that effective pressure at the glacier terminus is zero. The text will read:}
  The evolution of subglacial channel size $S_g$ is given as
  \begin{linenomath*}
    \begin{equation}
      \label{eq:dS_dt}
      \frac{\partial S_g}{\partial t} = C_1 \frac{Q \Delta h}{l} - C_2 \left(h_{o}-\overline{h}\right)^n\,S_g,
    \end{equation}
  \end{linenomath*}
  \noindent where $t$ is time, $C_1= (1-\rho_wc_pc_t)\,\frac{\rho_wg}{\rho_iL}$ and $C_2=2A(\frac{\rho_wg}{n})^n$ are constants (values in Table~1), $g$ is the acceleration due to gravity, $Q$ is water discharge, $\overline{h}=\frac{1}{2}(h+h_p)$ is the mean hydraulic head in the channel with $h_p$ being the proglacial hydraulic head equal to zero, $l$, $h_{o}= \frac{\rho_i}{\rho_w} h_{ice}$ is the mean ice overburden pressure expressed in meter water equivalent ($\rho_w$ is density of water; $\rho_i$ is density of ice), and $n$ is Glen's n \citep[usually $n=3$; ][]{glen1955}.
  The first term on the equation's right side represents the channel opening by frictional heating, while the following term represents channel closure from ice deformation.

  \textit{
    The reviewer also correctly points out that water de-pressurizes underneath the glacier, in agreement with the other reviewer, we have included a comment that we assume pressurized flow across the glacier bed, as is common in subglacial hydrology models, although this may not always be the case \citep{perolo2018}.
  }

@@

\item  \textbf{And the model should come with some caveats, e. g., you could say that it is a simplistic and possibly unreliable model. To be fair, this is done in section 5.3, but that is too late.}

  \textit{The ``Methods'' section is largely meant to describe the model, as opposed to discuss its behavior or simplicity. In response to this comment, we will move the ``Experiment limitations'' to a section that directly after the ``Results'' but before the ``Discussion.''}

\item \textbf{  One of the comments made is that because the sub-aerial model has an algebraic relation between width and discharge (equation 9), the relationship between velocity and discharge is also algebraic; but this is simply a choice of the model. In reality, the width of a channel will also evolve over (long) time through processes of bank erosion, and the use of equation 9 properly only involves a long term time average of discharge, so it is misleading to use it in a time-specific way. Actually, this is admitted at line 370, along with other limitations in the models.}

  \textit{We appreciate these comments. With regard to the first point about channel width evolution in the subaerial model, this matter is confronted in Section 4.2 and 4.3 (now 5.1).
    Here we show that across a range of exponents and relationships between channel cross section and water discharge, subglacial systems are still more sensitive to water discharge variations that subaerial ones.
    This increased variability occurs across a range of $\alpha$ values that describe the relationship between channel width and water discharge as shown in Figure 6 (now 7).
    As a result, we believe that these result hold across as range of water discharge- channel width relationships, including ones where $\alpha$ evolves in time.
    The exception, shown in Section 4.3 (now 6.1), is that when water discharge evolves more slowly that channel size. Here, the behavior between subaerial and subglacial is very similar, because channel area in both channel type adjust to discharge variations.}

### 2025-02-03 revised_v1/response_editor.tex

\newcommand{\TODO}[1]{\textbf{TODO #1}}
\newcommand{\ian}[1]{{\textbf{\color{blue}Ian says:} \color{blue} #1} }
\newcommand{\leif}[1]{{\textbf{\color{red}Leif says:} \color{red} #1} }
\newcommand{\frederic}[1]{{\textbf{\color{red}Frederic says:} \color{red} #1} }
\newcommand{\aref}[1]{\textbf{Reference #1}}
\newcommand{\alpine}{\textit{ALPINE}\,}
\newcommand{\icesheet}{\textit{ICESHEET}\,}
\title{}
\author{}

\begin{document}
Dear Editor,

\vspace{.3cm}

We thank you for taking on this manuscript and coordinating the reviews. We found both reviews very helpful for improving the manuscripts.

\vspace{.2cm}

Following Reviewer 1, we have modified the first objective of the manuscript to describe the conditions where covariance between sediment transport capacity and water discharge could occur.
To develop this point further, we have conducted additional experiments with a series of smoothed hydrographs to better quantify the conditions when covariance could occur.
Additionally, we have reorganized and modified the text in response to their comments.
Lastly, at several points, in the review, it was mentioned that a point was already established.
It could be a substantial oversight of the literature on our part as authors.
However, we would be grateful if the reviewer would provide the specific references in these cases to help us fully address their comments.

\vspace{.2cm}

With respect to Reviewer 2, we believe that their specific comments about the model have been addressed by generalizing the equations in questions.
We understand their critique about the accumulation of water flow along the glacier's length.
However, we cannot find a rational that would suggest that these findings would not over shorter channels or in a more complex (i.e. spatially distributed) model.
To further highlight their concerns, we have moved the ``Experiment limitations'' discussion to its own section following the ``Results.''

\vspace{.2cm}
Again we are grateful for your assistance with the manuscript and hope that you find our responses compelling

\vspace{.5cm}

### 2025-02-03 Delaneyetal2024.tex
small changes

The first numerical experiment aims to quantify the sources of increased variability in the subglacial model outputs as they exhibit different seasonal evolutions and peaks (Figure~\ref{fig:model_outs}).
Variable relationships between model outputs and water discharge emerge for the subaerial and subglacial cases as channel size evolves (Figure~\ref{fig:model_outs} a, e).
Because subaerial channels have no history dependence, each water discharge value in the subaerial channel produces a unique water velocity, shear stress, and width-integrated shear stress (Section~\ref{sect:sub_mode}).
This characteristic results in a perfect rank correlation between the variables and water discharge (Figure~\ref{fig:Qw_vari}).
An inconsistent relationship between subglacial model outputs and water discharge persists across the range of parameters examined in the ensemble runs (see Section~\ref{sect:ensemble}; Figure S15).
\begin{figure}[hbt!]
  \centering
  \includegraphics[width=0.75\linewidth]{Fig2.png}
  \caption{Model outputs from simulations using the hydrographs in panels a and e for the scenarios \alpine{} (a-d) and \icesheet{} (e-h).
    Black (gray) lines in a and e represent the subglacial channel size (water discharge).
    Blue (orange) lines represent outputs from the subglacial (subaerial) channel.
    Data are shown at $15$\unit{min} intervals.
    Arrows show examples where variables peak in the subglacial channel before the subaerial one.
    Insets in f-h show the peak melt event denoted by the shaded area in panels e--h, with 
  }
  \label{fig:model_outs}
\end{figure}

@@

\begin{acknowledgements}
  SNSF Project No. $\mathrm{PZ00P2}$\_$202024$ provided  funding for I. Delaney.
  A. Tedstone acknowledges funding from the European Research Council, award $818994$ -- CASSANDRA.
  G. King and M. J. Gevers provided useful comments on a previous version of this manuscript.
  We benefited from fruitful discussions with J. Irving.
  We thank two reviewers for constructive and critical comments on a previous version of the manuscript.
\end{acknowledgements}
\bibliographystyle{copernicus}
\bibliography{Paperlib.bib}

>>>>

The first numerical experiment aims to quantify the timing and covariance of the subglacial model outputs with respect to water discharge in response to different seasonal evolutions and peaks (Figure~\ref{fig:model_outs}).
Variable relationships between model outputs and water discharge emerge for the subaerial and subglacial cases as channel size evolves (Figure~\ref{fig:model_outs} a, e).
Because subaerial channels have no history dependence, each water discharge value in the subaerial channel produces a unique water velocity, shear stress, and width-integrated shear stress (Section~\ref{sect:sub_mode}).
This characteristic results in a perfect rank correlation between the variables and water discharge (Figure~\ref{fig:Qw_vari}).
An inconsistent relationship between subglacial model outputs and water discharge persists across the range of parameters examined in the ensemble runs (see Section~\ref{sect:ensemble}; Figure S15).
\begin{figure}[hbt!]
  \centering
  \includegraphics[width=0.75\linewidth]{Fig2.png}
  \caption{Model outputs from simulations using the hydrographs in panels a and e for the scenarios \alpine{} (a-d) and \icesheet{} (e-h).
    Black (gray) lines in a and e represent the subglacial channel size (water discharge).
    Blue (orange) lines represent outputs from the subglacial (subaerial) channel.
    Data are shown at $15$\unit{min} intervals.
    Arrows show examples where variables peak in the subglacial channel before the subaerial one.
    Insets in f-h show the peak melt event denoted by the shaded area in panels e--h, with 
  }
  \label{fig:model_outs}
\end{figure}

@@

\begin{acknowledgements}
  SNSF Project No. $\mathrm{PZ00P2}$\_$202024$ provided  funding for I. Delaney.
  A. Tedstone acknowledges funding from the European Research Council, award $818994$ -- CASSANDRA.
  G. King and M. J. Gevers provided useful comments on a previous version of this manuscript.
  We benefited from fruitful discussions with J. Irving.
  We thank B. Hubbard and an anonymous reviewer for constructive and critical comments on the manuscript.
\end{acknowledgements}
\bibliographystyle{copernicus}
\bibliography{Paperlib.bib}
